---
title: "covid 19"
author: "Kimberly Orantes"
date: "October 11, 2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

````{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

The first graph shows there is a higher number of females and a low number of males in this study. However, the number of males that died is greater than the amount of the females that died. It would be expected that if it were indiscriminate, then we would see the same amount of deaths for males and females, meaning sex wouldn't matter.    
  
  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The first table, first line tells us that out of all the people that remain alive 56% of them are females and 44% of them are males. The second line tells us that of all the people who are dead 36% are females and 64% are males.  
  
## Inferential results 




## Conclusions

